\documentclass{article}

\usepackage{graphicx}
\usepackage{amssymb,amsmath}
\usepackage{theorem}

\author{Nic Hollingum - 308193415}
\title{Complexity}

\begin{document}

\begin{definition}
Graph $G$
\begin{align}
	\nonumber G = & (V,E) \\
	\nonumber V = & \{v_1, v_2, ..., v_n\} \quad & (vertices)\\
	\nonumber E \subseteq & V^2 & (edges)
\end{align}
\end{definition}

\begin{definition}
Path $\pi$ for graph $G = (V,E)$ of length $l$
\begin{align}
	\nonumber \pi(v_a, v_b) = \{(x_i, x_{i+1}) : 0 \leq i < l \wedge x_i, x_{i+1} \in V \wedge x_0 = v_a \wedge x_l = v_b\}
\end{align}
\end{definition}

\begin{definition}
Pathset $\Pi_G$ for graph $G = (V,E)$
\begin{align}
	\nonumber \Pi_G(v_a, v_b) = \{\pi(v_a, v_b) : \pi(v_a, v_b) \subseteq E\}
\end{align}
Hence if there is no path from $v_a$ to $v_b$ in $G$: $\Pi_G(v_a, v_b) = \emptyset$
\end{definition}


\section{Multiterminal Cut Problem}

\begin{definition}
Multi-Terminal Cut

Given a graph $G=(V,E,w,T)$, a set $T=\{t_1, t_2, ... t_k\}$ of $k$ specified vertices or {\em terminals} : $T \subseteq V$ , and a positive weight $w(e)$ for each edge, find a minimum weight set of edges $E' \subseteq E$ such that the removal of $E'$ from $E$ disconnects each terminal from all others.
An instance of MTC is therefore defined fromally:

\begin{align}
	\nonumber MTC = & (V,E,w,T)\\
	\nonumber & E \subseteq V^2\\
	\nonumber & w : E \rightarrow \mathbb{Z}^+\\
	\nonumber & T \subseteq V
\end{align}

Solution Space: $2^E$

Solution Cost: $C : 2^E \rightarrow \mathbb{Z}^+$
\begin{align}
	\nonumber C(S) = \displaystyle\sum\limits_{(u,v) \in S} w(u,v)
\end{align}

Feasible Solution Space: $F \subseteq 2^E$
\begin{align}
	\nonumber \Pi_{(V,E \setminus S)}(t_i, t_j) = \emptyset \quad \forall (t_i, t_j) \in T^2 \setminus \{(t, t) | t \in T\}
\end{align}

Alternatively:

\begin{align}
	\nonumber \forall S \in F : S = \{(u, v) | \forall t \in T \quad & \Pi_{(V,E \setminus S)}(u, t) \neq \emptyset \\
	\nonumber & \quad \Rightarrow \Pi_{(V,E \setminus S)}(v, t) = \emptyset \quad or \\
	\nonumber & \Pi_{(V,E \setminus S)}(v, t) \neq \emptyset \\
	\nonumber & \quad \Rightarrow \Pi_{(V,E \setminus S)}(u, t) = \emptyset \}
\end{align}

Optimal Solution Space: $O \subseteq F$
\begin{align}
	\forall S_j \in F , \quad \forall S_i \in O \quad \nonumber C(S_i) \leq C(S_j)
\end{align}

\end{definition}


\section{Robust Actor Allocation Problem}

\begin{definition}
Given a graph $G=(V,E,w_c,w_i)$ and a set of processors $P$, where $w_c : E \rightarrow \mathbb{Z}^+$ and $w_i : V \times P \rightarrow \mathbb{Z}^+$, find a mapping $\alpha : V \rightarrow P$ such that:
$\displaystyle\sum\limits_{(u,v) \in E : \alpha(u) \neq \alpha(v)} w_c(u,v) + \displaystyle\sum\limits_{v \in V} w_i(v, \alpha(v))$ is minimised.

\begin{align}
	\nonumber RAAP = & (V,E,w_c, w_i, P)\\
	\nonumber & E \subseteq V^2\\
	\nonumber & w_c : E \rightarrow \mathbb{Z}^+\\
	\nonumber & w_i : V \times P \rightarrow \mathbb{Z}^+\\
	\nonumber & P = \{p_1, p_2, ...p_p\}
\end{align}

Solution Space: $\alpha \in P^V \subseteq P(V \times P)$

Solution Cost: $C : P^V \rightarrow \mathbb{Z}^+$
\begin{align}
	\nonumber C(\alpha) = \displaystyle\sum\limits_{(u,v) \in E : \alpha(u) \neq \alpha(v)} w_c(u,v) + \displaystyle\sum\limits_{v \in V} w_i(v, \alpha(v))
\end{align}

Feasible Solution Space: $F \subseteq P^V$
\begin{align}
	\nonumber C(\alpha) \neq \infty
\end{align}

Optimal Solution Space: $O \subseteq F$
\begin{align}
	\nonumber 	C(\alpha_i) \leq C(\alpha_j) \quad \forall \alpha_j \in F \quad \forall \alpha_i \in O
\end{align}

\end{definition}

\section{Mapping}

\begin{definition}
Mapping

\begin{align}
	\nonumber I \in & MTC(V,E,w,T) &\\
	\nonumber I' \in & RAAP(V', E', w_c, w_i, P) &\\
	\nonumber & I \mapsto I' if: &\\
	\nonumber & V' = V\\
	\nonumber & E' = E\\
	\nonumber & P = T\\
	\nonumber & w_c = w\\
	\nonumber & w_i(v,p) = 0 \forall v \in V', p \in P
\end{align}
\end{definition}

\begin{definition}
{\em Simplified RAAP}

To keep the proof simple we restrict RAAP to the SRAAP, which differs in that certain processors are pre-assigned.
This simplification is only applicale to the mapping, it imposes a restriction on the mapping.

We have required an isomorphism between terminals and processors in the mapping, this simplification makes that concrete.
Vertices which are terminals are restricted to being assigned to the processor which also represent that terminal.

We can trivially solve SRAAP using RAAP by making the cost of not assigning the pre-assigned vertices to their processors infinite.

Main Processor: $\mu : V \rightarrow P$
\begin{align}
	\nonumber \forall v_i \in P : \mu(v_i) = p_i
\end{align}

Solution Space: $\alpha \in P^V \subseteq P(V \times P)$

Feasible Solution Space: $F \subseteq P^V$
\begin{align}
	\nonumber C(\alpha) \neq \infty \\
	\nonumber \forall v_i \in P : \alpha(v_i) = \mu(v_i)
\end{align}

Optimal Solution Space: $O \subseteq F$
\begin{align}
	\nonumber 	C(\alpha_i) \leq C(\alpha_j) \quad \forall \alpha_j \in F \quad \forall \alpha_i \in O
\end{align}

\end{definition}

\begin{definition}
Reversal Function $\gamma$

\begin{align}
	\nonumber \gamma & : (P^V) \rightarrow P(E) \\
	\nonumber \gamma & \mapsto \{(u,v) \in E | \alpha(u) \neq \alpha(v)\}
\end{align}
\end{definition}

\begin{lemma}
\label{PATHASSIGN}
Any vertices with a path to one another in the residual graph are assigned to the same processor.

$\forall v_a, v_b \in V : \forall \pi \in \Pi_{(V, E \setminus \gamma(\alpha))}(v_a, v_b) : \forall (x_i, x_j) \in \pi : \alpha(x_i) = \alpha(x_j)$
\end{lemma}

\begin{proof}
This derives from the definition of $\gamma$.
We prove it by contradeiction:

presume $\exists (x_i, x_j) \in \pi : \alpha(x_i) \neq \alpha(x_j)$
\begin{align}
	\nonumber (x_i, x_j) & \in \gamma(\alpha) \\
	\nonumber (x_j, x_j) & \not\in E \setminus \gamma(\alpha) \\
	\nonumber \Rightarrow (x_i, x_j) & \not\in \pi
\end{align}

\end{proof}

\begin{lemma}
The mapping is always feasible

$\forall \alpha \in F' : \quad \gamma(\alpha) \in F$
\end{lemma}

\begin{proof}
We use $F'$ to denote the feasible solution space of (S)RAAP to distinguish it from that of MTC.

We show by contradiction that there can exist no feasible $\alpha$ for which $\gamma (\alpha)$ is not feasible for MTC.

Presume $\exists \beta \in F' : \gamma(\beta) \not\in F$, Then:

\begin{align}
	\nonumber \exists (u,v) \in \gamma(\beta) : \exists t \in T, u, v \in V s.t. & \quad \Pi_G (u, t) \neq \emptyset \wedge \Pi_G (v, t) \neq \emptyset \\
	\nonumber & \exists \pi_u \in \Pi_G (u, t) \Rightarrow \beta(u) = \beta(t) \quad \quad \mbox{(from \ref{PATHASSIGN})}\\
	\nonumber & \exists \pi_v \in \Pi_G (v, t) \Rightarrow \beta(v) = \beta(t) \\
	\nonumber & \Rightarrow \beta(u) = \beta(v) \\
	\nonumber & \Rightarrow (u,v) \not\in \gamma(\beta)
\end{align}

In plain english: if there was a solution which when reversed was infeasible, then there has to be a path from 2 endpoints of an element of the cutset in the residual graph.
But from \ref{PATHASSIGN} that means those 2 vertices are assigned to the same processor, which menas that edge couldnt have been in the cutset, which is a contradiction.
\end{proof}

\bibliographystyle{plain}
\bibliography{biblio}

\end{document}